---
title: 'DATA607 - Assignment Week 7'
subtitle: 'Working with XML and JSON in R'
author: "Donald Butler"
date: "10/10/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more
than one author. For each book, include the title, authors, and two or three other attributes that you find
interesting.

Create file representations of the selected books in the formats: HTML, XML, and JSON.

Are the three data frames identical?

## Load required R Libraries

```{r library, warning=FALSE, message=FALSE}
library(tidyverse)
library(XML)
library(jsonlite)
library(xml2)
library(rvest)
```

# Read HTML Data

```{r}
html_file <- 'http://raw.githubusercontent.com/dab31415/DATA607/main/Homework/Assignment_5/books.html'

html_df <- read_html(html_file) %>%
  html_elements('table') %>%
  html_table()

html_df
```
# Read XML Data

```{r}
xml_file <- 'http://raw.githubusercontent.com/dab31415/DATA607/main/Homework/Assignment_5/books.xml'

xml_df <- read_xml(xml_file) %>%
  xmlParse() %>%
  xmlToDataFrame()

xml_df

```

# Read JSON Data

```{r}
json_file <- 'https://raw.githubusercontent.com/dab31415/DATA607/main/Homework/Assignment_5/books.json'
json_df <- fromJSON(json_file)

json_df
```


