---
title: "DATA607 - Week 3 Assignment"
author: "Donald Butler"
date: "9/12/2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, warning=FALSE, message=FALSE}
library(tidyverse)
```

## Exercise 1

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
url = 'https://raw.githubusercontent.com/dab31415/DATA607/main/Homework/Assignment_3/majors-list.csv'
majors <- read_csv(url,show_col_types = FALSE)
majors %>%
  filter(grepl('DATA|STATISTICS',Major))

```

## Exercise 2

Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"  

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
raw_input <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'

expected_result <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```

Remove new line characters and positional indicators and trim whitespace.

```{r}
new_input <- raw_input %>%
  str_replace_all('\\n','') %>%
  str_replace_all('(\\[\\d+\\])','') %>%
  str_trim()

new_input
```

Introduce delimiter by searching for whitespace between double quotes, then remove the leading and trailing double quote.

```{r}
new_input <- new_input %>%
  str_replace_all('\\"[ ]*\\"',',') %>%
  str_replace('^"','') %>%
  str_replace('"$','')

new_input

```
 Split the input string by the delimiter and convert to a vector.

```{r}
new_input <- new_input %>%
  str_split(pattern = ',') %>%
  unlist()

new_input
```
 
Compare the converted input to our expected result.

```{r}
identical(expected_result,new_input)
```
## Exercise 3

Describe, in words, what these expressions will match:

```
(.)\1\1  
```
  Will match any character repeated 3 consecutive times, for example 'aaa'.

```
(.)(.)\2\1  
```
  Will match any two characters that are repeated reversed, for example 'abba'.

```
(..)\1  
```
  Will match any two characters that are repeated, for example 'mama'

```
(.).\1.\1  
```
  Will match any character repeated two times, but separated by a character in between each occurrence, for example 'anana' in 'banana',

```
(.)(.)(.).*\3\2\1  
```
  Will match any three characters that are later reversed with any number of characters between them. For example, 'abccba', 'abcdddcba'

## Exercise 4
Construct regular expressions to match words that:

Start and end with the same character.  
```
^(.).*\1$
```

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)  
```
(..).*\1
```

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)  
I had to escape '\' the astrics in the following regex to prevent R Markdown from converting to italics.
```
(.).*\1.*\1
```
