---
title: 'DATA607 - Assignment Week 10'
subtitle: 'Sentiment Analysis'
author: "Donald Butler"
date: "10/31/2021"
urlcolor: blue
bibliography: references.bib
biblio-style: apalike
link-citations: true
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

We were asked to recreate the R code from chapter 2 of the book "Text Mining with R: A Tidy Approach" then perform similar analysis on another authors set of works. 

## Load Packages

Load the packages required to complete the sentiment analysis.

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidytext)
library(janeaustenr)
library(wordcloud)
library(reshape2)
```

# Recreate Sentiment Analysis

Recreate the sentiment analysis from Chapter 2 in "Text Mining with R: A Tidy Approach". [@silge17]

## Sentiment Lexicons

A sentiment lexicon is a set of words and a sentiment score that is attributed to each word.

### AFINN-111

The AFINN sentiment lexicon scores words on a scale of -5 (negative sentiment) to 5 (positive sentiment). [@IMM2011-06010]

```{r}
get_sentiments("afinn")
```

### Bing

The Bing sentiment lexicon assigns words as either positive or negative. [@Hu04]

```{r}
get_sentiments("bing")
```

### NRC Word-Emotion Association Lexicon

The nrc sentiment lexicon classifies words as positive or negative, but also by emotions such as fear or anger. [@mohammad13]

```{r}
get_sentiments("nrc")
```

## Sentiment in Jane Austen's Books

Using the Bing sentiment lexicon to compare the sentiment of all of Jane Austen's books.

```{r}
lines_per_index <- 80

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text,regex("^chapter [\\divxlc]", ignore_case = TRUE)))
  ) %>%
  ungroup() %>%
  unnest_tokens(word,text)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments('bing'), by = 'word') %>%
  count(book, index = linenumber %/% lines_per_index, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sentiment = positive - negative)

jane_austen_sentiment %>%
  ggplot(aes(index, sentiment, fill = book)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~book, ncol = 2, scales = "free_x")

```

## Comparing Sentiment Lexicons

In Jane Austen's book "Pride and Prejudice", we will compare the three sentiment lexicons.

```{r}
pride_prejudice <- tidy_books %>%
  filter(book == "Pride & Prejudice")

afinn <- pride_prejudice %>%
  inner_join(get_sentiments('afinn'), by = 'word') %>%
  group_by(index = linenumber %/% lines_per_index) %>%
  summarize(sentiment = sum(value)) %>%
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>%
    inner_join(get_sentiments('bing'), by = 'word') %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>%
    inner_join(get_sentiments('nrc') %>%
                 filter(sentiment %in% c('positive','negative')), by = 'word') %>%
    mutate(method = 'NRC')
) %>%
  count(method, index = linenumber %/% lines_per_index, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sentiment = positive - negative)

bind_rows(afinn, bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~method, ncol = 1, scales = 'free_y')

```

## Common Positive and Negative Words

Determine the most common positive and negative words in Jane Austen's books.

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments('bing'), by = 'word') %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~sentiment, scales = 'free_y') + 
  labs(x = 'Contribution to sentiment', y = NULL)
```

## Custom Stop Words

The word 'miss' is contributing significantly towards negative sentiment, but it's use within Jane Austen's books isn't used in that context. Create a custom stop word list to adjust for anomalies in the sentiment lexicon used.

```{r}
custom_stop_words <- bind_rows(tibble(word = c('miss'), lexicon = c('custom')), stop_words)

custom_stop_words
```

## Wordclouds

Construct a wordcloud of the most common words in Jane Austen's books.

```{r}
tidy_books %>%
  anti_join(stop_words, by = 'word') %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

### Wordcoulds by Sentiment

Construct a wordcloud of the positive and negative words in Jane Austen's books.

```{r}
tidy_books %>%
  inner_join(get_sentiments('bing'), by = 'word') %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = 'n', fill = 0) %>%
  comparison.cloud(colors = c('gray20', 'gray80'), max.words = 100)
```

## Sentiment based on other units

Instead of considering sentiment on a word for word basis, we can look at it based on other units of text such as chapters. Determine the most negative chapter in each book. 

```{r}
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = 'regex', pattern = "Chapter|CHAPTER|chapter [\\divxlc]") %>%
  ungroup()

austen_chapters %>%
  group_by(book) %>%
  summarize(chapters = n())

bing_negative <- get_sentiments('bing') %>%
  filter(sentiment == 'negative')

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bing_negative, by = 'word') %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c('book','chapter')) %>%
  mutate(ratio = negativewords / words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>%
  ungroup()

```

# Sir Arthur Conan Doyle

Perform a similar analysis based on the works of author Sir Arthur Conan Doyle. 

## Project Gutenberg

Project Gutenberg is a library of public domain works. The gutenbergr package allows access to these works. [@gutenbergr] Retrieve a list of the Sherlock Holmes books that are available.

```{r}
library(gutenbergr)

doyle_books <- gutenberg_works(author == 'Doyle, Arthur Conan') %>%
  filter(gutenberg_bookshelf == 'Detective Fiction')

doyle_books
```

With the list of books, download the works and tidy. 

```{r}
tidy_doyle <- doyle_books %>%
  gutenberg_download(meta_fields = 'title') %>%
  group_by(gutenberg_id) %>%
  mutate(linenumber = row_number()) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

Unfortunately, in this collection of works, chapter pages are not present, so we are unable to consider a units by chapter.

## Sentiment in Sir Arthur Conan Doyal's Books

Using the Bing sentiment lexicon to compare the sentiment of all of the Sherlock Holmes books.

```{r, fig.height = 10}
lines_per_index <- 80

doyle_sentiment <- tidy_doyle %>%
  inner_join(get_sentiments('bing'), by = 'word') %>%
  count(title, index = linenumber %/% lines_per_index, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sentiment = positive - negative)

doyle_sentiment %>%
  ggplot(aes(index, sentiment, fill = title)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~title, ncol = 3, scales = "free_x")

```

## Comparing Sentiment Lexicons

In "A Study in Scarlet", we will compare the three sentiment lexicons.

```{r}
study_scarlet <- tidy_doyle %>%
  filter(title == "A Study in Scarlet")

scarlet_afinn <- study_scarlet %>%
  inner_join(get_sentiments('afinn'), by = 'word') %>%
  group_by(index = linenumber %/% lines_per_index) %>%
  summarize(sentiment = sum(value)) %>%
  mutate(method = "AFINN")

scarlet_bing_and_nrc <- bind_rows(
  study_scarlet %>%
    inner_join(get_sentiments('bing'), by = 'word') %>%
    mutate(method = "Bing et al."),
  study_scarlet %>%
    inner_join(get_sentiments('nrc') %>%
                 filter(sentiment %in% c('positive','negative')), by = 'word') %>%
    mutate(method = 'NRC')
) %>%
  count(method, index = linenumber %/% lines_per_index, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sentiment = positive - negative)

bind_rows(scarlet_afinn, scarlet_bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~method, ncol = 1, scales = 'free_y')

```

## Syuzhet 

I tried playing with the Syuzhet package to use another sentiment lexicon. [@syuzhet] The package computes a sentiment score for each line of the book text. 

```{r}
library(syuzhet)

study_in_scarlet <- doyle_books %>%
  filter(title == 'A Study in Scarlet') %>%
  gutenberg_download(meta_fields = 'title') %>%
  filter(grepl('[A-Za-z]',text))

syuzhet_sentiment <- get_sentiment(study_in_scarlet$text, method = 'syuzhet')
summary(syuzhet_sentiment)
```


## Common Positive and Negative Words

Determine the most common positive and negative words in the Sherlock Holmes books.

```{r}
doyle_bing_word_counts <- tidy_doyle %>%
  inner_join(get_sentiments('bing'), by = 'word') %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

doyle_bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~sentiment, scales = 'free_y') + 
  labs(x = 'Contribution to sentiment', y = NULL)
```

## Custom Stop Words

Again we see the word 'miss' on the common negative list, although not as impacting with Jane Austen's work, but based on the time period I would expect a fair number of instances are referring to a women and not the negative context.


## Wordclouds

Construct a wordcloud of the most common words in the Sherlock Holmes books.

```{r}
tidy_doyle %>%
  anti_join(stop_words, by = 'word') %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

### Wordcoulds by Sentiment

Construct a wordcloud of the positive and negative words in Jane Austen's books.

```{r}
tidy_doyle %>%
  inner_join(get_sentiments('bing'), by = 'word') %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = 'n', fill = 0) %>%
  comparison.cloud(colors = c('gray20', 'gray80'), max.words = 100)
```

## Sentiment based on other units

Unfortunately the chapter pages that were in Jane Austen's books, do not appear in the files for Sir Arthur Conan Doyle, so a similar analysis cannot be performed.

# References